Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-08-06 manuscripts/reply/reviews/response.tex

% \makeatletter
% \renewcommand{\section}{\@startsection{section}{1}{0mm}%
%     {-8pt}%
%     {4pt}%
% {\sffamily\large\itshape}}
% \makeatother

\begin{document}
% \raggedright
\setlength{\parindent}{0in}

I greatly appreciate the helpful comments of the editor and reviewers. In this
document, I provide responses to these comments, including descriptions of the
changes I have made.

\section{Editor's Comments}\label{ed}
\setcounter{commentCounter}{0}
\revcomment{
    The most negative review suggests that the authors (and Hickerson et al)
    might have reached a point of diminishing returns, where further exchange
    in the pages of Evolution might not be informative for our readership. My
    own view is somewhat in line with this reviewer. I do think there are some
    new ideas in the present manuscript, and I also want to honor the previous
    AE who solicited an additional response from Oaks and colleagues to the
    last Hickerson et al. paper. But I would also say that after this paper,
    enough is enough, in my opinion \ldots
}
\response{
    We agree. We are very appreciative of this opportunity to clarify some
    aspects of Hickerson et al.'s response, and feel our reply will make a nice
    end to an interesting discussion about Bayesian model selection.

    We have re-written a lot of the manuscript and feel the new version will be
    of general interest to readers interested in Bayesian selection of
    evolutionary models.
}

\revcomment{
     I also echo the suggestion from reviewer 2 to shorten this paper as much
     as is practical, especially by moving things to the supplement. One can
     safely assume that readers still following the thread have read the
     preceding papers in the series, I think.
}
\response{
    We are greatful for this insight by the editor and reviewers that our
    paper was too long and lacked focus.
    We greatly reduced the length of the main text, moved some items (including
    2 figures) to the supplement, and re-wrote a lot of the manuscript to focus
    on the most interesting aspects of Bayesian model choice brought to
    light by this series of papers.
}

@@

\revcomment{
    One alternative suggestion could be to formulate the problem as a
    hierarchical model, in which one seeks only to infer mean and variance in
    divergence times. Such a model would be much easier to fit, as the summary
    statistics ordering problem discussed (and solved) in the original msBayes
    paper could be circumvented easily, as the model would not seek to infer
    taxon specific divergence times. Also, it could then be compared to a model
    of zero variance using either Bayes factors, or by assessing the posterior
    probability of the variance to be zero using a simulation / bootstrap
    approach. If a model of co-divergence is preferred, the inferred posterior
    on the mean divergence would directly serve as an estimate of the joint
    divergence time. If independent divergences are inferred, then a simpler
    model of a single taxon pair could be used to infer the divergence times of
    each taxon pair individually, as they were found to be independent. For
    such a model, the same ABC simulations could be reused for all pairs,
    making this step computationally very cheap. But I'm sure there would also
    be other alternatives.
}
\response{
    These are great points raised by Reviewer 2. Many of these are
    discussed in detail in a separate paper \citep{Oaks2014dpp}.
    We highlight the improved methodology presented there to add a positive
    note to the response.
}

\revcomment{
    A minor issue on L53: while it is true that all numerical methods suffer
    from large prior ranges, it is also true that effect is pronounced in ABC,
    as the sampling occurs from the prior, and not the posteriors, as is done
    in techniques such as MCMC. Hence, it is not a problem that is completely
    extrinsic to the ABC machinery.
}
\response{
    We thank Reviewer 2 for pointing out this section, which was admitedly very
    confusing. We have re-written this section (and a lot of the manuscript) to
    try to make this distinction clearer. Our goal is to distinguish two
    different ways that broad uniform priors could cause the poor behavior of
    msBayes that was shown in the original Oaks et al. paper.

    We now pose these as alternative hypotheses to explain the methods tendency
    to support clustered divergence times. We also discuss the distinct
    predictions made by each hypothesis, and how best to improve the method
    under each of them.

    This has greatly improved the clarity, focus, and general interest of the
    paper, and we are very thankful that Reviewer 2 steered us in this
    direction.
}

### 2014-08-06 manuscripts/reply/reply-abstract.tex
More work on reply ms and review response.

Establishing that a set of population splitting events occurred at the same
time can be a potentially persuasive argument that the populations were
affected by the same geographic event.
Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
model choice method (\msb) to estimate such patterns of divergences across
taxa, to which \citet{Hickerson2013} responded.
Both papers agree that the primary inference enabled by the method is very
sensitive to prior assumptions and often erroneously supports shared
divergences across taxa when prior uncertainty about divergence times
is represented by a uniform distribution.
However, the papers differ in the best explanation and solution for this
problem.
\citet{Hickerson2013} suggested the lack of robustness in \msb analyses is
caused by numerical approximation error due to the inefficiency of the method's
rejection algorithm under broad uniform priors on divergence times
(Hypothesis~\ref{hypError}).
To overcome this, they proposed a model-averaging approach that uses narrow,
empirically informed uniform priors.
Here we demonstrate the approach of \citet{Hickerson2013} will often
erroneously support shared divergence events and is dangerous in the sense that
the empirically-derived uniform priors often exclude from consideration the
true values of the divergence-time parameters.
\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
marginal likelihoods to the uniformly distributed priors on divergence times
(Hypothesis~\ref{hypML}).
To mitigate this sensitivity, they suggested alternative probability
distributions to accommodate prior uncertainty about divergence times.
As predicted by fundamental principles of Bayesian model choice, more flexible
distributions can accommodate prior uncertainty about parameters without
placing excessive prior weight in vast regions of parameter space with low
likelihood;
this increases the method's robustness and power to detect temporal variation
in divergences \citep{Oaks2014dpp}.
We also show that the tendency of \msb analyses to support models of shared
divergences is primarily due to Hypothesis~\ref{hypML}, whereas
Hypothesis~\ref{hypError} is not a tenable explanation for the bias.
All of these papers have highlighted the importance of choosing prior
assumptions about parameters in Bayesian model selection, because they can
strongly influence the posterior probabilities of the models we seek to
estimate.
No amount of computation can rescue our inference if we choose priors on
parameters such that the exact posterior supports the wrong model of
evolutionary history.

>>>>

Establishing that a set of population splitting events occurred at the same
time can be a potentially persuasive argument that the populations were
affected by the same geographic event.
Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
model choice method (\msb) to estimate such a pattern of simultaneous
divergence across taxa, to which \citet{Hickerson2013} responded.
Both papers agree that the primary inference enabled by the method is very
sensitive to prior assumptions and often erroneously supports shared
divergences across taxa when prior uncertainty about divergence times
is represented by a uniform distribution.
However, the papers differ in the best explanation and solution for this
problem.
\citet{Hickerson2013} suggested the lack of robustness in \msb analyses is
caused by numerical approximation error due to the inefficiency of the method's
rejection algorithm under broad uniform priors on divergence times
(Hypothesis~\ref{hypError}).
To overcome this, they proposed a model-averaging approach that uses narrow,
empirically informed uniform priors.
\citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
marginal likelihoods to the uniformly distributed priors on divergence times
(Hypothesis~\ref{hypML}).
To mitigate this sensitivity, they suggested alternative probability
distributions to accommodate prior uncertainty about divergence times.
Here we use analyses of simulated and empirical data to show that the tendency
of \msb analyses to support models of shared divergences is primarily due to
Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
explanation for the bias.
We also demonstrate the approach of \citet{Hickerson2013} will often
erroneously support shared divergence events and is dangerous in the sense that
the empirically-derived uniform priors often exclude from consideration the
true values of the divergence-time parameters.
\citep{Oaks2014dpp} showed, as predicted by fundamental principles of Bayesian
model choice, more flexible distributions can accommodate prior uncertainty
about parameters without placing excessive prior weight in vast regions of
parameter space with low likelihood; this increases the method's robustness and
power to detect temporal variation in divergences.
All of these papers have highlighted the importance of choosing prior
assumptions about parameters in Bayesian model selection, because they
can strongly influence the posterior probabilities of the models we seek
to estimate.
If we choose priors placing too much weight in unlikely regions of parameter
space such that the true posterior supports the wrong model of evolutionary
history, no amount of computation can rescue our inference.
